---
title: "task2"
author: "Yanjiao"
date: "2022-11-26"
output: html_document
---

```{r setup, include=FALSE}
library(lambda.r)
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data}
imagem1 = read.table("imagem1.txt")
imagem2 = read.table("imagem2.txt")
imagem3 = read.table("imagem3.txt")

# change colnames
names = c("y", "x", "label", "NDAI", "SD", "CORR", "DF", "CF", "BF", "AF", "AN")
colnames(imagem1) = names
colnames(imagem2) = names
colnames(imagem3) = names
```

# Split image 1 using systematic assignment:
```{r}
summary(imagem1$y)
```

```{r}
summary(imagem1$x)
```

```{r}
imagem1$y
```


We first use the systematic assignment to split each image. The pattern for the split with $k=4$ is shown in the figure where k is the number of folds in one image. We justify the validity of the systematic assignment as it ensures lower dissimilarity between folds. As for the $3k$ folds of all the three images, we randomly select $2k$ folds as the training set, $\lceil 0.5k \rceil$ folds as the validation set and $\lfloor 0.5k \rfloor$ folds as the test set, which approximately yields a $4:1:1$ allocation for the three sets.

```{r}
# Referefence: https://onlinelibrary.wiley.com/doi/full/10.1111/ecog.02881
systematic.split = function(image, k){

  # prepare for the split
  image = image %>%
    mutate(level.x = as.factor(cut(image$x, breaks = 2*k)),
           level.y = as.factor(cut(image$y, breaks = 2*k)),)
  levels(image$level.x)=as.character(1:(2*k))
  levels(image$level.y)=as.character(1:(2*k))

  image = image %>%
    mutate(
      mod = (as.integer(level.x) + as.integer(level.y)) %% k) %>%
    mutate(fold = mod+1)
  
  return(image)
}
```

```{r}
imagem1.syst = systematic.split(imagem1, k=4)
imagem2.syst = systematic.split(imagem2, k=4)
imagem3.syst = systematic.split(imagem3, k=4)

# plot the split of image1
imagem1.syst %>%
  ggplot(aes(x = x, y = y, col = factor(fold)))+
  geom_point()+
  labs(x = "x", y = "y", color = "Fold")+
  scale_color_brewer(palette = "GnBu")+
  theme_bw()
  #scale_color_grey(start = 0.8, end = 0.2)
```












