---
title: "task2"
author: "Yanjiao"
date: "2022-11-26"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data}
imagem1 = read.table("imagem1.txt")
imagem2 = read.table("imagem2.txt")
imagem3 = read.table("imagem3.txt")

# change colnames
names = c("y", "x", "label", "NDAI", "SD", "CORR", "DF", "CF", "BF", "AF", "AN")
colnames(imagem1) = names
colnames(imagem2) = names
colnames(imagem3) = names
```

# Split image 1 using systematic assignment:
```{r}
summary(imagem1$y)
```

```{r}
summary(imagem1$x)
```

```{r include=FALSE}
imagem1$y
```

We first use the systematic assignment to split each image. The pattern for the split in image 1 is shown in the figure. The systematic assignment is a valid split as it ensures lower dissimilarity between folds. Among all the 12 blocks of the three images, 2 blocks are selected as the test set. As for the remaining 10 blocks, one block is used for validation and the rest for training when the number of folds $k$ is not greater than 10. If $k$ is larger than 10 (which is not a common case), after each of the 10 blocks has been used for validation, we randomly select one block to construct the additive $k-10$ folds. If $k=12$, for instance, after all the 10 blocks have been in the validation set in turn, we randomly select one block for validation and use the remaining 9 blocks for the training set and repeat this process again such that there are 12 cross validation sets in total. This paper uses 10-fold cross validation to assess the fit of models in the next section. We claim that resampling from the 10 blocks when more than 10 folds are needed generally is reasonable because the number of folds is usually between 5 and 10. [Maybe more justification later]


Referefence: https://onlinelibrary.wiley.com/doi/full/10.1111/ecog.02881

```{r}
systematic.split = function(image, i){
  
  # divide the image into 2T*2T pieces
  T = 4
  image = image %>%
    mutate(level.x = as.factor(cut(image$x, breaks = 2*T)),
           level.y = as.factor(cut(image$y, breaks = 2*T)),)
  levels(image$level.x)=as.character(1:(2*T))
  levels(image$level.y)=as.character(1:(2*T))

  image = image %>%
    mutate(
      mod = (as.integer(level.x) + as.integer(level.y)) %% T) %>%
    mutate(fold = mod+4*i-3) 
  
  return(image %>% select(-c(level.x, level.y, mod)))
}
```

```{r}
imagem1.syst = systematic.split(imagem1, i=1)
imagem2.syst = systematic.split(imagem2, i=2)
imagem3.syst = systematic.split(imagem3, i=3)

# combine the three images
image.syst = rbind(imagem1.syst, imagem2.syst, imagem3.syst)

# show the number of data in each fold
table(image.syst$fold)
```

```{r}
# create the test data
image.syst.test = NULL
set.seed(111111)

idx.test = c()
for (i in 1:2){
  idx = sample(1:length(unique(image.syst$fold)), size = 1)
  image.syst.test = rbind(image.syst.test, image.syst[image.syst$fold==idx, ])
  idx.test = c(idx.test, idx)
}

# create training data
image.syst.train = image.syst[!(image.syst$fold %in% idx.test), ]

table(image.syst.test$fold)
table(image.syst.train$fold)
```

```{r}
# plot the split of image1
imagem1.syst %>%
  ggplot(aes(x = x, y = y, col = factor(fold)))+
  geom_point()+
  labs(x = "x", y = "y", color = "Fold")+
  scale_color_brewer(palette = "GnBu")+
  theme_bw()
  #scale_color_grey(start = 0.8, end = 0.2)
```

# Split image 1 using buffering:

The second splitting method separates the training set from the validation set and the test set by creating a buffer. The cross in the middle of the image is used as the validation set. One of the four corners is randomly selected into the test set and the remaining three corners are in the training set. Although the pixels in the buffer area are not used as either the training set or the test set, the width of the pixels can be modified such that the fraction of the unused data can be controlled. One advantage of this split over the first one is that it generates spatially separated folds and ensures that no validation data or test data abuts the training data.

Reference (buffering): https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13107

```{r}
buf.split = function(image, width){
  xcut = quantile(image$x, c(4/9, 5/9, 1))
  ycut = quantile(image$y, c(4/9, 5/9, 1))
  
  image = image %>%
    mutate(
      level.x = case_when(
        ( (x < xcut[1]-width) | (x > xcut[2]+width) ) ~ 1,
        ( x > xcut[1] & x < xcut[2]) ~ 0,
        TRUE ~ 2),
      level.y = case_when(
        ( (y < ycut[1]-width) | (y > ycut[2]+width) ) ~ 1,
        ( y > ycut[1] & y < ycut[2]) ~ 0,
        TRUE ~ 2
          )
    )
  
  image = image %>%
    mutate(block = case_when(
      (level.x*level.y == 0) ~ 0,
      (level.x*level.y == 1) ~ 1,
      TRUE ~ 2
    ))
  return (image)
}
```

```{r}
imagem1.buf = buf.split(imagem1, width = 5)
imagem2.buf = buf.split(imagem2, width = 5)
imagem3.buf = buf.split(imagem3, width = 5)

# plot the split of image1
imagem1.buf %>%
  ggplot(aes(x = x, y = y, col = factor(block)))+
  geom_point()+
  labs(x = "x", y = "y", color = "Block")+
  scale_color_brewer(palette = "Blues")+
  theme_bw()
  #scale_color_grey(start = 0.8, end = 0.2)
```










